---
title: "Time series AirPassengers"
author: "Semiramis Castro-Jaimes"
date: "26/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Notes
The code follows the guide of the books 

- "Introductory Time Series with R" (2009) https://www.springer.com/gp/book/9780387886978

- "Introduction to Time Series Analysis and Forecasting" (2008). Link to the second edition (2015): https://www.wiley.com/en-us/Introduction+to+Time+Series+Analysis+and+Forecasting%2C+2nd+Edition-p-9781118745113


## Load data 
The dataset is available in base R
This dataset has be number of passengers (x1000) across time
```{r}
data(AirPassengers)
AP <- AirPassengers
class(AP)
```
## Explore data
```{r}
start(AP); end(AP); frequency(AP)
```
The data is anual; from 1949 to 1960

```{r}
summary(AP)
```

## Plot the data
```{r}
plot(AP, ylab = "Passengers (1000's)")
```

## Aggregate data anually
```{r}
plot(aggregate(AP))
```
We still see an increasing trend

## Plot the data by cycle
We capture the seasonal nature of the data
```{r}
boxplot(AP ~ cycle(AP))
```
We see that there is an increase in the number of passengers during the months June to September. This overlaps with the vacation period in summer

## Decompose the time series
```{r}
plot(decompose(AP, type="additive"))
```

```{r}
plot(decompose(AP, type="multiplicative"))
```

```{r}
AP_decom <- decompose(AP, type = "multiplicative")
str(AP_decom)
```
```{r}
Trend <- AP_decom$trend
Seasonal <- AP_decom$seasonal
ts.plot(cbind(Trend, Trend * Seasonal), lty = 1:2)
```

```{r}
plot(AP_decom$random)
```

```{r}
plot(stl(AP, s.window=2))

```

```{r}
str(stl(AP, s.window=2))
```
## Autocorrelation for raw data (only for exploration; not for analysis)
```{r}
acf(AP)
```

## Scatterplot of autocorrelations (lag 1)
```{r}
#length(AP)
plot(AP[1:143], AP[2:144])

```

## Random component plot (after decomposing into seasonality & trend)
```{r}
plot(AP_decom$random[7:138])
```

## Autocorrelation of the decomposed data (random component)
```{r}
acf(AP_decom$random[7:138])
```
If we square the highest peak (around 0.4), it says that the linear dependency of x(t) on x(t+1) would only explain 0.16% of variance in x(t) 

## Holt-Winters smoothing
```{r}
AP.hw <- HoltWinters(AP, seasonal = "mult")
plot(AP.hw)
```
## Forecasting 4 years
```{r}
AP.predict <- predict(AP.hw, n.ahead = 4 * 12)
ts.plot(AP, AP.predict, lty = 1:2)
```

```{r}
print(AP.hw)
#print(AP.hw$beta)
#print(AP.hw$gamma)
```

