---
title: "Autoregressive models"
author: "Semiramis Castro-Jaimes"
date: "31/7/2020"
output: html_document
---

## Notes
The code follows the guide of the books 

- "Introductory Time Series with R" (2009) https://www.springer.com/gp/book/9780387886978

- "Introduction to Time Series Analysis and Forecasting" (2008). Link to the second edition (2015): https://www.wiley.com/en-us/Introduction+to+Time+Series+Analysis+and+Forecasting%2C+2nd+Edition-p-9781118745113


## Simulation of an auto-regressive (AR) process
The AR model assumes that the values can be predicted using solely the previous values (historical data); hence the name "autoregressive"
```{r}
set.seed(1)
x <- w <- rnorm(100)
for (t in 2:100) x[t] <- 0.7 * x[t - 1] + w[t]
plot(x, type = "l")

```

## Get the autocorrelation plot
```{r}
acf(x)
```
The decreasing sinusoidal pattern is typical of a stationary time series


## Partial autocorrelation plot
"The partial autocorrelation at lag k results after removing the effect of any correlations due to terms at shorter lags"
```{r}
pacf(x)
```
Here only the first lag is significative

